Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2024-04-21 doc/manuscript/submission_files/research_report_2023-22711R.tex
Create final files for submission

The mountain yellow-legged (MYL) frog, composed of the sister species
\emph{Rana muscosa} and \emph{Rana sierrae} \citep{vredenburg2007}, is
emblematic of the global declines of amphibians caused by Bd
\citep{scheele2019}. Once the most common amphibian in the high
elevation portion of California's Sierra Nevada mountains
\citep[USA,][]{grinnell1924}, during the past century this frog has
disappeared from more than 90\% of its historical range
\citep{vredenburg2007}. Due to the severity of its decline and the
increasing probability of extinction, both species are now listed as
``endangered'' under the U.S. Endangered Species Act. In the Sierra
Nevada, this decline was initiated by the introduction of non-native
trout into the extensive historically-fishless region
\citep{bradford1989, knapp2000} starting in the late 1800s. The arrival
of Bd in the mid-1900s and its subsequent spread \citep{vredenburg2019}
caused additional large-scale population extirpations
\citep{vredenburg2010, rachowicz2006}. These Bd-caused declines are
fundamentally different from the fish-caused declines because fish
eradication is feasible \citep{knapp1998} and results in the rapid
recovery of frog populations \citep{knapp2007, vredenburg2004}. In
contrast, Bd appears to persist in habitats even in the absence of
amphibian hosts \citep{walker2007}, and therefore represents a long-term
alteration of invaded ecosystems that amphibians will need to overcome
to reestablish populations.

@@

@@

To determine whether MYL frogs from recovering populations can be used
to reestablish extirpated populations, we conducted 24 reintroductions
in Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them to 1 of 12 nearby recipient
sites (Figure S1). The 3 donor populations are among
the largest MYL frog populations in Yosemite, and included 2 of the 5
recovering populations used in the frog evolution study referenced above
(see \textbf{Results - Frog evolution in response to Bd:} population 1
and 4). In addition, this subset of 2 donor populations provided frogs
for 20 of the 24 translocations. Following translocation, we estimated
adult survival and recruitment of new adults from capture-mark-recapture
(CMR) surveys and obtained counts of tadpoles and juveniles from visual
encounter surveys (VES). Across all translocation sites, the duration of
survey time series was 1--16 years (median = 5).

@@

@@

@@

Results from the frog translocation study indicated that most
populations showed evidence of successful reproduction and recruitment,
and that adult survival was often relatively high (described above).
Although suggestive of population establishment, a decade or more of
surveys may be necessary to confirm that populations are in fact
self-sustaining \citep{joseph2018}. To extend our inferences of
population establishment beyond those possible from the site-specific
CMR data, we developed a population viability model. Specifically, to
test whether the observed yearly adult survival probabilities in
translocated populations were sufficient for long-term viability, we
built a stage-structured matrix model that captured known frog
demography and included demographic and environmental stochasticity. We
parameterized the model using CMR data from translocated populations and
known life history values in this system (Table S1).

@@

To determine whether MYL frog populations show genomic patterns
consistent with an evolutionary response to Bd, we compared frog exomes
(i.e., coding region of a genome) between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in 4 populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in 5 populations that experienced a
Bd epizootic during the past several decades and have since recovered to
varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. Bd-exposure histories of the 9 study
populations are based on 10-20 years of VES and Bd surveillance using
skin swabbing \citep[e.g.,][]{knapp2016, zhou2015, wilber2022}. Naive
populations are characterized by large numbers of adults (i.e, typically
1000s), Bd prevalence that is generally 0\% except during occasional Bd
failed invasions \citep[during which Bd loads remain very
low,][]{wilber2022}, and no history of Bd epizootics since we first
surveyed these populations in the late 1990s and early 2000s
\citep{zhou2015}. In contrast, recovering populations exist in an
enzootic state \citep{briggs2010}, characterized by smaller numbers of
adults (generally \textless{} 500), high Bd prevalence \citep[often
\textgreater{} 80\%,][]{knapp2011}, and, in adults, moderate Bd loads
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}. Naive and recovering populations can be
identified unambiguously using these differences in Bd prevalence and
load. Finally, there is no potential for frog dispersal between the 9
study populations due to intervening distances and topography, as well
as the presence of introduced (predatory) fish and fish-induced habitat
fragmentation.

It is important to note that our genomic analysis is necessarily limited
by the small number of naive populations remaining on the landscape
(nearly all of which we sampled as part of the current study) and their
clumped distribution (Figure~\ref{fig-allelefrequencies}). These
characteristics of naive populations precluded a replicated analyses
among pairs of geographically proximate naive and recovering
populations. Instead, to obtain adequate power, we conducted a pooled
analysis across population types to identify genomic regions showing
patterns consistent with selection. Our comparative approach cannot
provide a direct assessment of the strength or timing of natural
selection, but does allow us to identify specific regions of the genome
that merit further study for their potential contribution to frog
resistance in recovering populations. (See \textbf{Supporting
Information - Frog evolution in response to Bd - Study design} for
additional details regarding the study design.)

@@

Results from the individual variant and splined window analyses show
that recovering populations show patterns consistent with selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S6B, C
SI). One of the 7 identified genes (LOC108802036) does not have an
associated annotation. For the outlier SNPs, frequency differences
between the naive and recovering populations ranged from 0.41 to 0.86.
Most of these SNPs showed frequency differences in only a subset of the
sampled populations (Figure~\ref{fig-allelefrequencies}A, B), but the
SNP in the RIN3 gene showed consistent differences in frequencies across
all populations (Figure~\ref{fig-allelefrequencies}C). This is
suggestive of parallel selection at this locus across multiple
populations. The other 6 outlier variants showed less consistent
frequency differences across the study populations, but for these we
still found statistically significant differences in 2 of the 3 genetic
clusters (containing populations 5--9;
Figure~\ref{fig-allelefrequencies}, Figure S6A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe results that
are consistent with parallel evolutionary changes following Bd exposure.

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 35 annotated genes were found in the 9 shared regions. Given
this large number of genes, here we focus on those with the strongest
signal of selection and/or immune-related functions. The largest
\(\pi_{diff}\), suggestive of directional selection, occurred in a 163kb
region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP
(Figure~\ref{fig-spline-manhattan}C). This region contains
approximately 500 SNPs and one annotated gene called
``interferon-induced very large GTPase 1-like'' (GVINP1). Additionally,
a shared outlier region on Contig1 contained two complement factor genes
(C6 and C7). Interestingly, this region had a large negative
\(\pi_{diff}\) value, consistent with balancing selection. Finally, one
shared outlier region on Contig8 contained one outlier SNP (TCF19) and
was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S7).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region is part of the extended MHC Class I and III
complex and shows synteny with other amphibian genomes.

Although the joint processes of Bd-caused population declines and
selection in response to Bd exposure could affect genetic diversity of
recovering populations, we found no consistent differences in
individual-level heterozygosity or population-level \(\pi\) between
naive and recovering populations (see \textbf{Supporting Information -
Frog evolution in response to Bd - Genetic diversity} for details).
Thus, despite patterns consistent with localized selection in particular
regions of the genome, we did not find evidence for reduced genetic
diversity across the genome in recovering populations. In addition, no
gene ontology (GO) biological functions, molecular functions, or
cellular processes were over-represented in either the outlier variants
or the 35 genes located in the overlapping \emph{F\textsubscript{ST}}
and \(\pi_{diff}\) splined windows (see \textbf{Supporting Information -
Frog evolution in response to Bd} \textbf{- GO analysis} for details).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but broadly-applicable strategies to recover
affected species are generally lacking \citep[e.g.,][]{garner2016}.
Here, we tested the possibility that populations of resistant
individuals from naturally recovering populations can be used to
reestablish extirpated populations of the endangered MYL frog in the
presence of a highly virulent fungal pathogen (Bd). Our results indicate
(i) the capacity of reintroduced populations to become established and
eventually recover despite ongoing disease, (ii) that 50\% of the
recovering populations are likely to persist over a 50-year period,
(iii) that there are genomic differences between naive and recovering
MYL frog populations, consistent with evolutionary change in frogs
following Bd exposure, and (iv) that some of the genomic differences are
in regions that contain genes related to disease resistance.
Collectively, these results (Figure~\ref{fig-recovery-model}) provide a
rare example of amphibian recovery in the presence of Bd, and have
important implications for the conservation and recovery of amphibians
and other taxa worldwide that are endangered by escalating impacts from
emerging infectious diseases. In light of the generally low success rate
of amphibian reintroduction efforts \citep{dodd2005}, our success in
reestablishing MYL frog populations via translocation of resistant
individuals is striking, and even more so given that MYL frogs were
driven to near-extinction by Bd.

@@

Importantly, some genomic regions that we identified as possibly under
selection are associated with cellular and immunological mechanisms
known to contribute to disease resistance, including in amphibians
\citep{zamudio2020}. For example, the MHC plays an important role in
immunity. In our study, we identified a region that shows evidence of
selection in recovering populations and contains eight genes associated
with either the MHC Class I or Class III regions. These results
corroborate numerous previous studies linking MHC genes to amphibian
resistance against Bd \citep[e.g.,][]{savage2011, bataille2015}.
Similarly, the region with the strongest pattern of possible directional
selection (as measured by \(\pi_{diff}\)) contains the
interferon-related gene GVINP1. Several previous studies of amphibians
have found this gene to be differentially expressed during Bd infection
\citep[e.g.,][]{grogan2018b, ellison2014} and in populations differing
in Bd susceptibility \citep{grogan2018b}. This gene is also strongly
linked to disease in salmon, explaining a notable 20\% of the resistance
phenotype \citep[i.e., gill damage and amoebic load as heritable host
resistance traits,][]{robledo2020, robledo2018}. We also identified a
region, characterized by high \emph{F\textsubscript{ST}} and low
\(\pi_{diff}\), that contained the complement genes C6 and C7. The
complement system plays an important role in innate immunity
\citep{rieraromo2016}, and our results could indicate that balancing
selection is acting in this region of the genome to favor a diverse set
of alleles, as is known for C6 in humans \citep{soejima2005}. Based on
the analysis of individual outlier variants, the RIN3 gene showed a
consistent pattern of allele frequency differences across all nine of
the frog populations sampled in this study, suggesting possible
consistent selection in populations distributed across a wide geographic
area. This gene is associated with immune response and in \emph{Xenopus}
is expressed during appendage regeneration \citep{fukazawa2009}.
Finally, the outlier variant with the lowest p-value was the
uncharacterized gene LOC108802036. In the genome of another frog
species, this gene is located adjacent to a type I interferon gene
(Np-IFNi2) \citep{gan2018}, and together with GVINP1 further suggests
the importance of interferon-related genes in this system. Collectively,
the genes associated with these genomic differences may confer at least
some degree of resistance against Bd infection, an attribute that may be
critically important to population reestablishment and recovery in the
presence of Bd.

Reintroduction of resistant MYL frogs was remarkably successful in
reestablishing viable populations in the presence of Bd. Of the 12
translocated populations, approximately 80\% showed evidence of both
successful reproduction and recruitment of new adults. Year-1 survival
for 12 of the 24 translocated cohorts exceeded 50\%, and \textgreater{}
70\% of translocated cohorts had survival above this 50\% level when the
earliest translocations are excluded (i.e., translocations conducted
when methods were still being refined; see \textbf{Materials and Methods
- Frog population recovery - Field methods} for a brief description of
these refinements). The fact that the relatively low Bd loads and
correspondingly high frog survival was maintained when frogs were moved
from donor populations to recipient sites indicates that these
characteristics of naturally-recovering populations were not solely an
effect of site characteristics, but were also strongly influenced by
resistance inherent in the frogs. Although it could be argued that the
relatively invariant Bd loads before versus after translocation are a
consequence of similar pathogen pressure in the donor and translocated
populations, this is at odds with the fact that in the first year after
translocation frog densities are typically 1-2 orders of magnitude lower
in the translocated versus donor populations and pathogen pressure
should follow a similar pattern. In addition to the maintenance of Bd
load and frog survival between natal and translocation sites, the
relatively high survival of translocated frogs was maintained in their
progeny, as expected if resistance has a genetic basis.

Results from the population viability model were also encouraging. In
particular, translocated populations with \textgreater{} 50\% survival
in the first year post-translocation were predicted to have a low
probability of extinction over 50 years (probability of extinction \(<\)
0.5 when year-1 juvenile survival probability was greater than 0.10).
The viability model highlighted the important role of frog survival in
affecting long-term population viability, and allowed us to extend the
temporal scale of our study beyond the years covered by our
post-translocation surveys. These long-term forecasts are important,
given that reintroduced MYL frog populations may often take decades to
achieve our ultimate goal of self-sustainability \citep{joseph2018}.
Making well-supported projections about the long-term outcome of
reintroduction efforts from shorter-term information is critically
important to the process of adaptive management of species
reintroduction programs \citep{seddon2007}, including the one we are
carrying out for MYL frogs. Specifically, the combined results from our
reintroduction study and viability modeling indicate that survival of
frogs in the first year following translocation is an effective proxy of
longer-term survival and population viability. In addition, given the
repeatability of frog survival at a site, 1-year frog survival also
serves as an effective proxy of site quality (i.e., the ability of a
site to support high frog survival and a viable frog population over the
long term). This proxy of site quality is important in the MYL frog
system because accurately predicting the ability of a site to support a
viable frog population \emph{a priori} remains difficult, even after
conducting 24 translocations over 16 years.

>>>>

The mountain yellow-legged (MYL) frog, composed of the sister species
\emph{Rana muscosa} and \emph{Rana sierrae} \citep{vredenburg2007}, is
emblematic of the global decline of amphibians caused by Bd
\citep{scheele2019}. Once the most common amphibian in the high
elevation portion of California's Sierra Nevada mountains
\citep[USA,][]{grinnell1924}, during the past century this frog has
disappeared from more than 90\% of its historical range
\citep{vredenburg2007}. Due to the severity of its decline and the
increasing probability of extinction, both species are now listed as
``endangered'' under the U.S. Endangered Species Act. In the Sierra
Nevada, this decline was initiated by the introduction of non-native
trout into the extensive historically-fishless region
\citep{bradford1989, knapp2000} starting in the late 1800s. The arrival
of Bd in the mid-1900s and its subsequent spread \citep{vredenburg2019}
caused additional large-scale population extirpations
\citep{vredenburg2010, rachowicz2006}. These Bd-caused declines are
fundamentally different from the fish-caused declines because fish
eradication is feasible \citep{knapp1998} and results in the rapid
recovery of frog populations \citep{knapp2007, vredenburg2004}. In
contrast, Bd appears to persist in habitats even in the absence of
amphibian hosts \citep{walker2007}, and therefore represents a long-term
alteration of invaded ecosystems that amphibians will need to overcome
to reestablish populations.

@@

\begin{figure*}
{\centering \includegraphics[width=7in]{conceptual_model_7in.pdf}}
\caption{\label{fig-recovery-model}For MYL frogs, a conceptual model
depicting the Bd-caused decline and subsequent natural recovery (black
text), facilitated recovery via reintroductions, and the linkages
between these two pathways. Rectangles and hexagons represent outcomes
and processes, respectively. Blue text indicates components that are
included in the current study. We quantified frog survival and
population viability following reintroductions directly using field
studies and models, and evaluated ``disease-mediated selection''
indirectly using a comparison of frog exomes in naive and recovering
populations. The timeline shows the general sequence of the components,
with the dashed portion indicating a projection into the future.}
\end{figure*}

@@

To determine whether MYL frogs from recovering populations can be used
to reestablish extirpated populations, we conducted 24 reintroductions
in Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them to 1 of 12 nearby recipient
sites (Figure S1). The 3 donor populations are among
the largest MYL frog populations in Yosemite, and included 2 of the 5
recovering populations used in the frog evolution study referenced above
(see \textbf{Results - Frog genome evolution:} populations 1 and 4).
These 2 donor populations provided frogs for 20 of the 24
translocations. Following translocation, we estimated adult survival and
recruitment of new adults from capture-mark-recapture (CMR) surveys and
obtained counts of tadpoles and juveniles from visual encounter surveys
(VES). Across all translocation sites, the duration of survey time
series was 1--16 years (median = 5).

@@

\begin{figure}
{\centering \includegraphics[width=3.4in]{translocation_survival_bysiteid_3pt42in.pdf}}
\caption{\label{fig-translocation-survival}Median 1-year survival for
each cohort of translocated frogs at the 12 recipient sites, as
estimated for each site from the mrmr CMR model. Error bars show the
95\% uncertainty intervals. Sites are arranged along the x-axis using
the average of the median 1-year survival per translocation at each
site. Dot colors indicate the donor population from which frogs in each
translocated cohort were collected. When multiple translocations were
conducted to a site, points and error bars are slightly offset to avoid
overlap.}
\end{figure}

@@

\begin{SCfigure*}[\sidecaptionrelwidth][t!]
{\centering \includegraphics[width=5.5in]{conditional_effects_plot_4pt5in.pdf}}
\caption{\label{fig-cond-effects}Results from the among-site
meta-analysis showing conditional effects of the important predictors of
1-year frog survival (expressed as a probability): (A) winter severity
in the year following translocation, (B) elevation of recipient site,
(C) donor population, and (D) sex. In (A) and (B), blue lines are
medians and gray ribbons are 95\% uncertainty intervals. In (C) and (D),
box plots show medians, first and third quartiles, largest and smallest
values within 1.5x interquartile range, and values outside the 1.5x
interquartile range.}
\end{SCfigure*}

@@

Results from the frog translocation study are suggestive of population
establishment. However, a decade or more of surveys may be necessary to
confirm that populations are in fact self-sustaining \citep{joseph2018}.
To extend our inferences of population establishment beyond those
possible from the site-specific CMR data, we developed a population
viability model. Specifically, to test whether the observed yearly adult
survival probabilities in translocated populations were sufficient for
long-term viability, we built a stage-structured matrix model that
captured known frog demography and included demographic and
environmental stochasticity. We parameterized the model using CMR data
from translocated populations and known life history values in this
system (Table S1).

@@

To determine whether MYL frog populations show genomic patterns
consistent with an evolutionary response to Bd, we compared frog exomes
(i.e., coding region of a genome) between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in 4 populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in 5 populations that experienced a
Bd epizootic during the past several decades and have since recovered to
varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. Bd-exposure histories of the 9 study
populations are based on 10-20 years of VES and Bd surveillance using
skin swabbing \citep[e.g.,][]{knapp2016, zhou2015, wilber2022}. Naive
populations are characterized by large numbers of adults (i.e, typically
1000s), Bd prevalence that is generally 0\% except during occasional Bd
failed invasions \citep[during which Bd loads remain very
low,][]{wilber2022}, and no history of Bd epizootics since we first
surveyed these populations in the late 1990s and early 2000s
\citep{zhou2015}. In contrast, recovering populations exist in an
enzootic state \citep{briggs2010}, characterized by smaller numbers of
adults (generally \textless{} 500), high Bd prevalence \citep[often
\textgreater{} 80\%,][]{knapp2011}, and, in adults, moderate Bd loads
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}. Naive and recovering populations can be
identified unambiguously using these differences in Bd prevalence and
load. There is no potential for frog dispersal between the 9 study
populations or any other populations due to intervening distances and
topography, as well as the presence of introduced (predatory) fish and
fish-induced habitat fragmentation.

\begin{figure*}
{\centering \includegraphics[width=7in]{pop_viability_figures_for_manuscript_7in.pdf}}
\caption{\label{fig-viability}Results from population viability
analysis: (A) Predicted long-run growth rate \(\lambda\) for different
values of yearly adult survival probability \(\sigma_{A_R}\) and year-1
juvenile survival and recruitment probability \(\sigma_{J_1}\), given
the parameterized, deterministic model. Colored points show the
predicted \(\lambda\) values for the twelve translocated populations
when year-1 juvenile survival probability is \(\sigma_{J_1} = 0.09\)
(indicated by the dashed line). The red line shows where
\(\lambda = 1\). Note that the point for 70413 is mostly hidden behind
other points. (B) Predicted 50-year extinction probabilities of the 12
translocated populations, given demographic stochasticity, environmental
variability in \(\sigma_{J_1}\), and different mean values of
\(\sigma_{J_1}\). There are 6 lines at extinction probability = 1, 5 of
which (70414--70619) are hidden beneath the line for 70505 when
\(\sigma_{J_1} < 0.10\). (C) 100 simulated trajectories (black lines)
from the population viability model that most closely matched the
observed abundance trajectory of adult amphibians at site 70550 (light
orange).}
\end{figure*}

It is important to note that our genomic analysis is necessarily limited
by the small number of naive populations remaining on the landscape
(nearly all of which we sampled as part of the current study) and their
clumped distribution (Figure~\ref{fig-allelefrequencies}). These
characteristics of naive populations precluded a replicated analysis
among pairs of geographically proximate naive and recovering
populations. Instead, to obtain adequate power, we conducted a pooled
analysis across population types to identify genomic regions showing
patterns consistent with selection. Our comparative approach cannot
provide a direct assessment of the strength or timing of natural
selection, but does allow us to identify specific regions of the genome
that merit further study for their potential contribution to frog
resistance in recovering populations. (See \textbf{Supporting
Information - Frog genome evolution - Study design} for additional
details.)

@@

Results from the individual variant and splined window analyses show
that recovering populations show patterns consistent with selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S6B, C). One of the 7 identified genes (LOC108802036) does not have an
associated annotation. For the outlier SNPs, frequency differences
between the naive and recovering populations ranged from 0.41 to 0.86.
Most of these SNPs showed frequency differences in only a subset of the
sampled populations (Figure~\ref{fig-allelefrequencies}A, B), but the
SNP in the RIN3 gene showed consistent differences in frequencies across
all populations (Figure~\ref{fig-allelefrequencies}C). This is
suggestive of parallel selection at this locus across multiple
populations. The other 6 outlier variants showed less consistent
frequency differences across the study populations, but for these we
still found statistically significant differences in 2 of the 3 genetic
clusters (containing populations 5--9;
Figure~\ref{fig-allelefrequencies}, Figure S6A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe results that
are consistent with parallel evolutionary changes.

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 35 annotated genes were found in the 9 shared regions. Given
this large number of genes, here we focus on those with the strongest
signal of selection and/or immune-related functions. The largest
\(\pi_{diff}\), suggestive of directional selection, occurred in a 163kb
region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP
(Figure~\ref{fig-spline-manhattan}C). This region contains
approximately 500 SNPs and one annotated gene called
``interferon-induced very large GTPase 1-like'' (GVINP1). Additionally,
a shared outlier region on Contig1 contained two complement factor genes
(C6 and C7). Interestingly, this region had a large negative
\(\pi_{diff}\) value, consistent with balancing selection. Finally, one
shared outlier region on Contig8 contained one outlier SNP (TCF19) and
was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S7).
This region (854kb from the beginning of the outlier region to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region is part of the extended MHC Class I and III
complex and shows synteny with other amphibian genomes. To provide
details specifically on possible selection in the Yosemite populations
described in \textbf{Results - Frog population recovery}, \(\pi\) values
for the 9 shared outlier regions are shown in
Figure S8 for each of the four Yosemite
populations; two of these populations provided frogs for 20 of the 24
frog translocations.

\begin{SCfigure*}[\sidecaptionrelwidth][t!]
{\centering \includegraphics[width=4.5in]{allele_maps_4pt5in.pdf}}
\caption{\label{fig-allelefrequencies}Genomic patterns consistent with
selection on individual variants in recovering MYL frog populations at
the landscape scale (based on a linear mixed model analysis). For each
of the 9 naive and recovering MYL frog populations (indicated by
numbered points), adjacent pie charts show allele frequencies for the 11
outlier SNPs from 7 distinct genes: (A) LOC108802036, (B) TCF19, VARS,
MAP1S, TYRP1, and SRFBP1, and (C) RIN3. Charts are superimposed on a map
of the Sierra Nevada study area, with Yosemite, Kings Canyon, and
Sequoia National Parks (from north to south) shown in dark gray, and the
range boundary for MYL frogs shown in light gray. The inset map locates
the study area and range boundary in California. Sites 1 and 4 (site id
= 72996 and 70567, respectively) were also used as sources of frogs for
reintroductions described in \textbf{Results - Frog population
recovery}.}
\end{SCfigure*}

Although the joint processes of Bd-caused population declines and
selection in response to Bd exposure could affect genetic diversity of
recovering populations, we found no consistent differences in
individual-level heterozygosity or population-level \(\pi\) between
naive and recovering populations (see \textbf{Supporting Information -
Frog genomic evolution - Genetic diversity} for details). Thus, despite
patterns consistent with localized selection in particular regions of
the genome, we did not find evidence for reduced genetic diversity
across the genome in recovering populations. In addition, no gene
ontology (GO) biological functions, molecular functions, or cellular
processes were over-represented in either the outlier variants or the 35
genes located in the overlapping \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) splined windows (see \textbf{Supporting Information -
Frog genomic evolution} \textbf{- GO analysis} for details).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but broadly-applicable strategies to recover
affected species are generally lacking \citep[e.g.,][]{garner2016}.
Here, we tested the possibility that populations of resistant
individuals from naturally recovering populations can be used to
reestablish extirpated populations of the endangered MYL frog in the
presence of a highly virulent pathogen (Bd). Our results indicate (i)
the capacity of reintroduced populations to become established and
eventually recover despite ongoing disease, (ii) that 50\% of the
reintroduced populations are likely to persist over a 50-year period,
(iii) that there are genomic differences between naive and recovering
MYL frog populations, consistent with evolutionary change in frogs
following Bd exposure, and (iv) that some of the genomic differences are
in regions that contain genes related to disease resistance.
Collectively, these results (Figure~\ref{fig-recovery-model}) provide a
rare example of amphibian recovery in the presence of Bd, and have
important implications for the conservation and recovery of amphibians
and other taxa worldwide that are endangered by escalating impacts from
emerging infectious diseases. In light of the generally low success rate
of amphibian reintroduction efforts \citep{dodd2005}, our success in
reestablishing MYL frog populations via reintroduction of resistant
individuals is striking, and even more so given that MYL frogs were
driven to near-extinction by Bd.

\begin{SCfigure*}[\sidecaptionrelwidth][t!]
{\centering \includegraphics[width=4.5in]{splinewindow_manhattan_4pt5in.pdf}}
\caption{\label{fig-spline-manhattan}Results from the splined window
analysis showing patterns consistent with selection on genomic regions
in recovering MYL frog populations. Note that several key candidate
genes (RIN3, TCF19, VARS) are outliers in both the individual variant
analysis (Figure~\ref{fig-allelefrequencies}) and splined window
analysis (this figure). Manhattan plots show outlier regions for the
difference in (A) nucleotide diversity \(\pi_{diff}\) and (B)
\emph{F\textsubscript{ST}}. In (A), outlier regions are shown above the
upper red dashed line and below the lower red dashed line. In (B),
outlier regions are shown above the single dashed red line. Outlier
regions for either \(\pi_{diff}\) or \emph{F\textsubscript{ST}} are
shown in blue and outlier regions for both \(\pi_{diff}\) and
\emph{F\textsubscript{ST}} are shown in red. (C) Magnified Contig19 from
(A) showing two adjacent outlier regions for \(\pi_{diff}\) 12.9Mb
upstream of the RIN3 outlier SNP (indicated with a dashed vertical blue
line). (D) Magnified Contig8 from from (B) showing the
\emph{F\textsubscript{ST}} outlier region that includes the outlier SNPs
TCF19 and VARS. This region of the genome contains 8 annotated genes
known to occur in the extended MHC Class I and III regions.}
\end{SCfigure*}

@@

Importantly, some genomic regions that we identified as possibly under
selection are associated with cellular and immunological mechanisms
known to contribute to disease resistance, including in amphibians
\citep{zamudio2020}. For example, the MHC plays an important role in
immunity. We identified a region containing eight genes associated with
the MHC Class I and III regions that showed patterns consistent with
selection in recovering populations. These results align with numerous
previous studies linking MHC genes to amphibian resistance against Bd
\citep[e.g.,][]{savage2011, bataille2015}. Similarly, the region with
the strongest pattern of possible directional selection (as measured by
\(\pi_{diff}\)) contains the interferon-related gene GVINP1. Several
previous studies of amphibians have found this gene to be differentially
expressed during Bd infection \citep[e.g.,][]{grogan2018b, ellison2014}
and in populations differing in Bd susceptibility \citep{grogan2018b}.
This gene is also strongly linked to disease in salmon, explaining a
notable 20\% of the resistance phenotype \citep[i.e., gill damage and
amoebic load as heritable host resistance
traits,][]{robledo2020, robledo2018}. We also identified a region,
characterized by high \emph{F\textsubscript{ST}} and low \(\pi_{diff}\),
that contained the complement genes C6 and C7. The complement system
plays an important role in innate immunity \citep{rieraromo2016}, and
our results could indicate that balancing selection is acting in this
region of the genome to favor a diverse set of alleles, as is known for
C6 in humans \citep{soejima2005}. Based on the analysis of individual
outlier variants, the RIN3 gene showed a consistent pattern of allele
frequency differences across all nine of the frog populations sampled in
this study, suggesting possible consistent selection in populations
distributed across a wide geographic area. This gene is associated with
immune response and in \emph{Xenopus} is expressed during appendage
regeneration \citep{fukazawa2009}. Finally, the outlier variant with the
lowest p-value was the uncharacterized gene LOC108802036. In the genome
of another frog species, this gene is located adjacent to a type I
interferon gene (Np-IFNi2) \citep{gan2018}, and together with GVINP1
further suggests the importance of interferon-related genes in this
system. Collectively, the genes associated with these genomic
differences may confer at least some degree of resistance against Bd
infection, an attribute that may be critically important to population
reestablishment and recovery in the presence of Bd.

Reintroduction of resistant MYL frogs was remarkably successful in
reestablishing viable populations in the presence of Bd. Of the 12
translocated populations, approximately 80\% showed evidence of both
successful reproduction and recruitment of new adults. Year-1 survival
for 12 of the 24 translocated cohorts exceeded 50\%, and \textgreater{}
70\% of translocated cohorts had survival above this 50\% level when the
earliest translocations are excluded (i.e., translocations conducted
when methods were still being refined; see \textbf{Materials and Methods
- Frog population recovery - Field methods} for a brief description of
these refinements). The fact that the relatively low Bd loads and
correspondingly high frog survival was maintained when frogs were moved
from donor populations to recipient sites indicates that these
characteristics of naturally-recovering populations were not solely an
effect of site characteristics, but were also strongly influenced by
intrinsic characteristics of frogs, including resistance. Although it
could be argued that the relatively invariant Bd loads before versus
after translocation are a consequence of similar pathogen pressure in
the donor and reintroduced populations, this is at odds with the fact
that in the first year after translocation frog densities are typically
1-2 orders of magnitude lower in the reintroduced versus donor
populations and pathogen pressure should follow a similar pattern. In
addition to the maintenance of Bd load and frog survival between natal
and recipient sites, the relatively high survival of translocated frogs
was maintained in their progeny, as expected if resistance has a genetic
basis.

Results from the population viability model were also encouraging. In
particular, translocated populations with \textgreater{} 50\% survival
in the first year post-translocation were predicted to have a low
probability of extinction over 50 years (probability of extinction \(<\)
0.5 when year-1 juvenile survival probability was greater than 0.10).
The viability model highlighted the important role of frog survival in
affecting long-term population viability, and allowed us to extend the
temporal scale of our study beyond the years covered by our
post-translocation surveys. These long-term forecasts are important,
given that reintroduced MYL frog populations may often take decades to
achieve our ultimate goal of self-sustainability \citep{joseph2018}.
Making well-supported projections about the long-term outcome of
reintroduction efforts from shorter-term information is crucial for
adaptive management of species reintroduction programs
\citep{seddon2007}, including the one we are carrying out for MYL frogs.
Specifically, the combined results from our reintroduction study and
viability modeling indicate that survival of frogs in the first year
following translocation is an effective proxy of longer-term survival
and population viability. In addition, given the repeatability of frog
survival at a site, 1-year frog survival also serves as an effective
proxy of site quality (i.e., the ability of a site to support high frog
survival and a viable frog population over the long term). This proxy of
site quality is important in the MYL frog system because accurately
predicting the ability of a site to support a viable frog population
\emph{a priori} remains difficult, even after conducting 24
translocations over 16 years.
